---
title: "Benchmarking {SpeedyMarkov}"
author: "Sam Abbott"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking {SpeedyMarkov}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 7, 
  fig.align = "center"
)
```

## Benchmarking setup

The `SpeedyMarkov` and reference implementations have been benchmarked with a range of optimisations including making use of `Rcpp` and parallisation. Several parallisation approaches have been explored as has the use of a varying number of cores. The benchmarking grid has been implemented within `SpeedyMarkov` - see below for details.

* Load required packages.

```{r}
library(SpeedyMarkov)
library(microbenchmark)
library(parallel)
library(furrr)
library(ggplot2)
```

* Code for the benchmarking grid.

```{r}
SpeedyMarkov::benchmark_markov
```

All benchmarking results have been run elsewhere (`inst/scripts/benchmarking.R`) and then saved (`inst/benchmarks`). These benchmarks can be recreated locally by running the following in the package directory (*Warning:: This is likely to take upwards of an hour and a half*).

```{bash, eval = FALSE}
Rscript inst/scripts/benchmarking.R
```

## Two state Markov model

The two state Markov model has been benchmarked for both 100 cycles and 200 cycles (duration) with 100,000 samples each time - repeated 10 times per benchmark.

### Duration: 100

Benchmarking results show that both `Rcpp` implementations consistently outperform other implementations with the full `Rcpp` implementation being overall slightly faster but with potentially some instability. Parallisation appears to work well and scale with the number of cores used. Both parallisation approaches work equally well with `furrr` preferred due to its support for Windows. On a 4 core machine a parallel approach takes around 10% of the reference approaches runtime. There was also a substantial speed up - though less dramatic - when compared to the R `SpeedyMarkov` approach. 

Increasing the number of cycles (duration) resulted in a an approximate doubling of the runtime for both the reference and R `SpeedyMarkov` approaches. However, all C++ based approaches had only small increases in runtime highlighting the increased performance that C provides (though the difference between the partial and full C++ approaches did increase). In fact the parallel C++ implementations showed no apparent increase in runtime when the duration was doubled. This indicates that this approach should scale well to more complex models and/or longer runtimes.

* Table

```{r}
two_state_bench_dur_100 <- readRDS("../inst/benchmarks/two_state_duration_100.rds") 

two_state_bench_dur_100
```

* Plot

```{r, width = 16, height = 12}
ggplot2::autoplot(two_state_bench_dur_100) +
  theme_minimal()
```

### Duration: 200

* Table

```{r}
two_state_bench_dur_200 <- readRDS("../inst/benchmarks/two_state_duration_200.rds")

two_state_bench_dur_200
```


* Plot

```{r, width = 16, height = 12}
ggplot2::autoplot(two_state_bench_dur_200) +
  theme_minimal()
```
